---
title: "Quebec Elections the Effect of Quebecois Protectionism"
author: "Ben Draskovic, Arjun Dhatt, Gantavya Gupta"
date: "10/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
```

The Sampling method 

```{r}
set.seed(624)
Quebes_survey_data <- tibble(
  Gender = sample(x = c("Male", "Female", "Non-Binary", "Other", "Prefer not to say"), 
                  size = 100000, replace = TRUE, prob = c(0.48, 0.50, 0.01, 0.001, 0.009)), 
  #Based on Stats Can Census 2016 (Statistics Canada 2019)
  Age = sample(x = c("18 - 30 years old", "30 - 45 years old", "45 - 60 years old", "60+ years old", "17 years old or younger"), 
               size = 100000, replace = TRUE, prob = c(0.170, 0.192, 0.217, 0.252, 0.169)), 
  #Based on Stats Can Census 2016 (Statistics Canada 2019)
  Region = sample(x = c("Urban", "Town", "Rural"), size = 100000, replace = TRUE, prob = c(0.592, 0.086, 0.322)), 
  #Based on Duffin's 2016 surveys (Duffin 2019)
  Last_election = sample(x = c("I did not vote in the last provincial election", "I voted for the Liberal party", "I voted for the Conservative party", "I voted for the NDP party", "I voted for the Green party", "I voted for the Bloc Quebecois", "I voted for a party not listed above"), size = 100000, replace = TRUE, prob = c(0.327, 0.231, 0.108, 0.073, 0.030, 0.218, 0.013)), 
  #Based on Elections Canada results in 2019 (Elections Canada 2019)
  Upcoming_election = sample(x = c("I do not plan to vote", "Liberal Party", "Conservative Party", "NDP Party", "Green Party", "Bloc Quebecois", "A party not listed above"), size = 100000, replace = TRUE, prob = c(0.327, 0.23, 0.11, 0.024, 0.038, 0.26, 0.011)),
  #Took the difference in percentage of votes each party got from 2015-2019 in  Quebec. Divided that difference by 2 making the assumption that the extreme trends would slow and assumed the same level of voter participation (Elections Canada 2019)
  Area_for_Improvement = sample(x = c("Quebec Independenc", "Immigration", "Education", "Health Care", "Environment", "Other"), size = 100000, replace = TRUE, prob = c(0.15,0.15,0.1,0.3,0.15,0.15)),
  #The probabilities are more difficult to simmulate for this variable. However based upon the finding of Angus Reid Institute's Survey these are some of the stand out categories. Also the rough distribution found was a large focus on Health care with fairly similar rate of interest in other issues (Angus Reid Institute 2018)
  Protect_Culture = sample(x = c(1:5), size = 100000, replace = TRUE, prob = c(0.1, 0.2, 0.3, 0.2, 0.2)),
  #Based upon the the authors' perception that most in quebec belief that some level of protection is needed, and the continued existance of succession movements where 35% of Angus Reid Institute's survery respondants favored Quebec Sovereignty  (Angus Reid Institute 2018)
  Identity_Under_Threat = sample(x = c("Yes, Quebecois identity is under threat.", "No, it is a fictional issue.", "There are some threats but they are exaggerated.", "Don't Know"), size = 100000, replace = TRUE, prob = c(0.4,0.3,0.25,0.05))
  #Based upon the response to Angus Reid Institute's survey on Quebec Sovereignty, assuming a slightly higher rate would beleive that there is a credible threat than those seeking Independence, and that there would be a split among those not seeking independence on how they respond between Fictional issue and exagerated  (Angus Reid Institute 2016)
)

head(Quebes_survey_data)

```

# Non-Technical Executive Summary

# Introduction

# Survey Methodology

# Results

# Discussion
What the results could mean 

Weaknesses and what needs to come next 


#Appendices 
Where we put the Survey 




# References 
- Angus Reid Institute. (2016). What makes us Canadian? A study of values, beliefs, priorities and identity. Retrieved from http://angusreid.org/canada-values/
- Angus Reid Institute. (2018). Quebecers were more likely to say the province was on the wrong track than the right one before Oct. 1. Retrieved from http://angusreid.org/quebec-election-2018-analysis
- Duffin, P. B., & 29, A. (2019, April 29). Population distribution of Quebec, by rural/urban type 2016. Retrieved from https://www.statista.com/statistics/608694/population-distribution-of-quebec-by-rural-urban-type/#:~:text=This statistic shows the population,in large urban population centers.
- Elections Canada. (2019). General Election Results Raw Data. Retrieved from https://www.elections.ca/content.aspx?section=res&dir=rep/off/43gedata&document=index&lang=e
- Statistics Canada. (2019, August 09). Census Profile, 2016 Census Quebec [Province] and Canada [Country]. Retrieved from https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=24&Geo2=PR&Code2=01&Data=Count&SearchText=24&SearchType=Begins&SearchPR=01&B1=All&Custom=&TABID=3


